Concurrent, Parallel Garbage Collection in Linear Time

### 2016-05-27 finaletd/mainabstract.tex
abstract change?

\chapter*{Abstract}
\label{ch:abstract}
Garbage collection is slowly moving from being an utility program to requirement of every modern programming languages. Manual memory management offers great control in sequential programs. With multi-cores and distributed systems, most programs written recently are heavily multi-threaded and distributed. Generally, programs written that uses multi-threading and distributed are called concurrent programming. Concurrent programming are characterized by multiple independent processes/threads, communicate among them, uncertain order of concurrent operations among them.

The uncertainity in the order of operations makes the manual memory management of the concurrent programs difficult. Languages like C++ offers support to manage memory using smart pointers. These smart pointers are simple reiterated garbage collection algorithm working for simple object graph in heap. In fact, smart pointer usage does not guarantee you from memory leaks. For cyclic object graph, the concurrent process must identify that it is creating cycle and must use special type of smart pointer to indicate it is a cyclic object graph to help smart pointers collect garbage. Smart pointers require hints from process to avoid memory leaks. Garbage collectors can avoid all memory leaks, dangling pointers, and double deletion problems in any programming environment.

Concurrent programming comes in two flavors : shared memory and distributed memory. Shared memory systems helps threads to communicate with each other through shared memory. Most shared memory systems contain single garbage collector that runs concurrently. In distributed memory system, multiple systems contain process that communicate through messages sent across network. These systems require synchronization barrier among all the garbage collector in each system to solve the problem. Our thesis focuses on designing garbage collection algorithms for each environment that takes advantage of the environment.

This thesis offers solution to multi-collector shared memory garbage collection where multiple garbage collectors can work concurrently with application and among themselves to perform garbage collection. Multi-collector garbage collection in distributed systems requires synchronization among all the garbage collectors. This thesis aims at designing a multi-collector garbage collector that avoid global barrier among all garbage collectors in all active systems. Our garbage collection algorithms are designed to achieve solution that is localized, highly-scalable and complete.

>>>>

\chapter*{Abstract}
\label{ch:abstract}
Garbage collection is slowly moving from being a utility program to a requirement of every modern programming language. Manual memory management offers great control in sequential programs. With multi-cores and distributed systems, most programs written recently are heavily multi-threaded and distributed. Generally, programs were written that uses multi-threading and distributed are called concurrent programming. Concurrent programming is characterized by multiple independent processes/threads, communicate among them, the uncertain order of concurrent operations among them.

The uncertainty in the order of operations makes the manual memory management of the concurrent programs difficult. Languages like C++ offers support to manage memory using smart pointers. These smart pointers are simply reiterated garbage collection algorithm working for simple object graph in the heap. In fact, smart pointer usage does not guarantee you from memory leaks. For cyclic object graph, the concurrent process must identify that it is creating the cycle and must use the special type of smart pointer to indicate it is a cyclic object graph to help smart pointers collect garbage. Smart pointers require hints from process to avoid memory leaks. Garbage collectors can avoid all memory leaks, dangling pointers, and double deletion problems in any programming environment.

Concurrent programming comes in two flavors : shared memory and distributed memory. Shared memory systems help threads to communicate with each other through shared memory. Most shared memory systems contain single garbage collector that runs concurrently. In distributed memory system, multiple systems contain the process that communicates through messages sent across a network. These systems require synchronization barrier among all the garbage collector in each system to solve the problem. Our thesis focuses on designing garbage collection algorithms for each environment that takes advantage of the environment.

This thesis offers the solution to multi-collector shared memory garbage collection where multiple garbage collectors can work concurrently with an application and among themselves to perform garbage collection. Multi-collector garbage collection in distributed systems requires synchronization among all the garbage collectors. This thesis aims at designing a multi-collector garbage collector that avoid global barrier among all garbage collectors in all active systems. Our garbage collection algorithms are designed to achieve the solution that is localized, highly-scalable and complete.
